Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity

### 2022-10-14 papers/paper-1/sections/results.tex
Update results section

\drafting{
Figures to do
\begin{itemize}
\item Example of forward-modelled semimajor axis and eccentricity
\item Example of forward-modelled velocity dispersion profile
\end{itemize}
and maybe another one of BH merger remnant properties.
}

@@

\begin{figure*}
    \includegraphics[width=0.7\textwidth]{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.png}
    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution. Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs). The observed data is overlaid, and seen to lie well within the 50\% BCI. }}
    \label{fig:postpred}
\end{figure*}

@@

As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
For the AC progenitor class, we find the core radius to be $\sim 1.4\,\mathrm{kpc}$, effective radius $\sim 7.2\,\mathrm{kpc}$ (recall progenitor A had $R_\mathrm{e} \sim 4\,\mathrm{kpc}$), normalising density of $\sim 3.1\times10^9\,\Msun\,\mathrm{kpc}^{-2}$, core slope index of $\sim 0.05$, and S\'ersic index of $\sim 1.8$.

\drafting{
    \begin{itemize}
        \item corner plot description
        \item comparison between rperi values, maybe new section?
    \end{itemize}
}

@@

\begin{figure*}
    \includegraphics{compare_HH_gqs_res.png}
    \caption{\drafting{Comparison of core sersic fit for different mass resolutions. Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something. y-axes should read PDF}}
    \label{fig:cs_mass_compare}
\end{figure*}

>>>>

\drafting{
Figures to do
\begin{itemize}
\item Example of forward-modelled semimajor axis and eccentricity
\end{itemize}
and maybe another one of BH merger remnant properties.
}

@@

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.pdf}
    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution. Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs). The observed data is overlaid, and seen to lie well within the 50\% BCI. }}
    \label{fig:postpred}
\end{figure*}

\begin{figure*}
    \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats.}}
    \label{fig:cs_compare_naive}
\end{figure*}

@@

As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the mean and standard deviation of the sample. 
The mean value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates: this is seen most clearly for the latent parameter $n$, whose Gaussian error estimate implies that by $2\sigma$, there is a non-zero probability of $n$ assuming a value less than 0, which is unphysical.

We show the latent parameter distributions for each mass resolution family in \autoref{fig:cs_mass_compare}, and compare the median and IQRs \drafting{ensure this acronym explained somewhere} in \autoref{fig:cs_mass_compare_2}. 
We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
Secondly, we observe that not all latent parameters in the core-S\"ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution. \drafting{More particles --> core better resolved. Discuss here or in discussion?}

Whilst we are unable to attribute the error in \autoref{fig:cs_mass_compare} to a particular phenomenon, we may at least rule out that the spread is due to post-processing analysis introduced from differing observation projections of the merger remnant. 
We re-determine the mean and standard deviation of the logarithm of the projected density profile using a subset of the total 30 projections we have available to us for the H-1.00 merger family.
Repeating this data-reduction process four times, we then rerun the HMC algorithm for each subsample.
We compare the latent parameter distributions for each subsample to the latent parameter distributions for the H-1.00 family determined using the full projection sample, and find:
\begin{enumerate}
    \item The error in the latent parameters is the consistent when using a subsample of projections, or the full sample, implying
    \item The error in the latent parameters is independent of the projection direction.
\end{enumerate}
Consequently, we may attribute the uncertainty in the latent parameter distributions to one of:
\begin{itemize}
    \item Intrinsic scatter inherent to the physical system, or 
    \item Numerical noise arising from a source independent of mass resolution, or 
    \item A combination of the two.
\end{itemize}

\begin{figure*}
    \includegraphics{compare_HH_gqs_res.pdf}
    \caption{\drafting{Comparison of core sersic fit for different mass resolutions. Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something.}}
    \label{fig:cs_mass_compare}
\end{figure*}

\begin{figure*}
    \includegraphics{compare_HH_gqs_res_2d.pdf}
    \caption{\drafting{Same as \autoref{fig:cs_mass_compare}, but plotted as a function of mass resolution. Points correspond to the median estimate for the latent parameters, and error bars to the IQR.}}
    \label{fig:cs_mass_compare_2}
\end{figure*}

@@

### 2022-10-14 papers/paper-1/sections/stats.tex
Divide paper into two papers, better division of scientific message
Update results section

@@

In addition to visual inspection of the posterior predictive distribution, we wish to quantitatively assess the statistical ability of our model to capture the key features we are interested in.
Critically, the exact form of such a test will depend on which features we define as being `important' for the model to capture.
Typically though, a sense of how well a posterior distribution captures the first and second moments of a distribution (i.e., the mean and variance) are metrics we will be guided by.
\begin{enumerate}
    \item Does the posterior predictive distribution capture the \textit{location} of the data generating process? If our model predicts data $y^\mathrm{sim}$ significantly offset from our observed data $y$, then the model is not very useful in predicting future observations generated from the same data generating process as $y$.
    \item Does the posterior predictive distribution capture the \textit{scale} of the data generating process? \drafting{What does this imply??}
\end{enumerate}

>>>>

\section{Statistical Analysis}

@@

\drafting{Now using LOO method. Which section should this go in? Relation to HDI?}
In addition to visual inspection of the posterior predictive distribution, we wish to quantitatively assess the statistical ability of our model to capture the key features we are interested in.
Critically, the exact form of such a test will depend on which features we define as being `important' for the model to capture.
Typically though, a sense of how well a posterior distribution captures the first and second moments of a distribution (i.e., the mean and variance) are metrics we will be guided by.
\begin{enumerate}
    \item Does the posterior predictive distribution capture the \textit{location} of the data generating process? If our model predicts data $y^\mathrm{sim}$ significantly offset from our observed data $y$, then the model is not very useful in predicting future observations generated from the same data generating process as $y$.
    \item Does the posterior predictive distribution capture the \textit{scale} of the data generating process? \drafting{What does this imply??}
\end{enumerate}
